---
title: "Classifying using Logistic Regression"
author: "Salerno"
date: "2020-01-13"
categories: ["R Programming", "Classification", "Logistic Regression"]
tags: []
draft: FALSE
banner: img/banners/banner-5.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warnings = FALSE, echo = TRUE)
```

## 1 - Objective

The objective of this example is to identify each of a number of benign or malignant classes.

## 2 - Data

Let's getting the data.

```{r}

BCData <- read.table(url("https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data"), sep = ",")

```

```{r}

# setting column names
names(BCData)<- c('Id', 'ClumpThickness', 'CellSize','CellShape',     
                   'MarginalAdhesion','SECellSize', 'BareNuclei',
                   'BlandChromatin','NormalNucleoli',
                   'Mitoses','Class')

```

## 3 - EDA - Exploratory Data Analysis

It's important to extract prelimionary knowledge from the dataset.

```{r}

dim(BCData)

```

```{r}

str(BCData)

```

```{r}

# Remove rows with missing values
BCData<-BCData[!(BCData$BareNuclei=="?"),]

```

```{r}

# convert variable to an integer
BCData$BareNuclei<-as.integer(BCData$BareNuclei)

```

```{r}

str(BCData)

```

```{r}

summary(BCData)

```

```{r}

table(BCData$Class)

```

The number 2 means "benign" and number 4 means "malignant".


```{r}

boxplot(BCData[,2:10])

```

To better identify the presence of outliers, we can plot histograms of the variables in the database

```{r}

par(mfrow=c(3, 3))
hist(BCData$ClumpThickness)
hist(BCData$CellSize)
hist(BCData$CellShape)
hist(BCData$MarginalAdhesion)
hist(BCData$SECellSize)
hist(as.numeric(BCData$BareNuclei))
hist(BCData$BlandChromatin)
hist(BCData$NormalNucleoli)
hist(BCData$Mitoses)

```

## 4 - Model Fitting

```{r}

BCData$Class<-replace(BCData$Class,BCData$Class==2,0)
BCData$Class<-replace(BCData$Class,BCData$Class==4,1)

```

```{r}

table(BCData$Class)

```

It is time to build the logistic regression model to obtain a tool capable of predicting the class of new observations.

```{r}

LoGModel <- glm(Class ~.-Id, 
              family=binomial(link='logit'),data=BCData)

```

```{r}

summary(LoGModel)

```

```{r}

anova(LoGModel, test="Chisq")

```

```{r}

LGModelPred <- round(predict(LoGModel, type="response"))

```

```{r}

table(LGModelPred)

```

```{r}

table(BCData$Class, LGModelPred)

```

```{r}

library(caret)
confusionMatrix(table(LGModelPred, BCData$Class), positive="1") 

```

```{r}
library(pROC)

RocObj <- pROC::roc(BCData$Class, LGModelPred, levels = c(0, 1), direction = "<")

print(RocObj)

```

```{r}

plot.roc(RocObj)

```

```{r}

plot(RocObj, print.auc=TRUE, auc.polygon=TRUE, grid=c(0.1, 0.2),
               grid.col=c("green", "red"), max.auc.polygon=TRUE,
                       auc.polygon.col="blue", print.thres=TRUE)

```

