---
title: "Exponential Smoothing Model"
author: "Salerno"
date: '2019-12-18'
slug: exponential-smoothing-model
tags: ["exponential smoothing", "models"]
categories: models
draft: FALSE
banner: img/banners/banner-5.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warnings = FALSE, echo = TRUE)
```

## Introduction

The main idea here is breaking the ice in terms of exponential smoothing models

First of all it is importan to show some behaviours patterns usually found in time series

1) Trends: it is usually observed when the time series follow one specific direction. It can be linear or not.

2) Seasonality: it is a pattern that repeat in a certain times (specific period)

3) Cycle: Like seasonality but it appears in non specific time

```{r}
library(fpp)

data(elecequip)

plot(elecequip, xlab = "Time", ylab = "New Orders Index")

```

```{r}
decomp <- decompose(elecequip, type = "additive")
plot(decomp)
```

```{r}
seasonality_adjust <- elecequip - decomp$seasonal
plot(seasonality_adjust)
```

```{r}
library(mFilter)

hpfilter(elecequip, type = "lambda")
par(mfrow= c(2,1))
plot(elecequip, xlab = "Time", ylab = "New Orders Index")
lines(hpfilter(elecequip, type = "lambda")$trend, col = "red", lwd = 2)
legend(1996, 200, c("Original Serie", "Trend - HP Filter"), col = c("black", "red"), lwd = c(1,2), bty = "n")
plot(hpfilter(elecequip, type = "lambda")$cycle,  xlab = "Time", ylab = "New Orders Index")

```

```{r}
data("cafe")

hpfilter(cafe, type = "lambda")
par(mfrow= c(2,1))
plot(cafe, xlab = "Time", ylab = "Expenditures Quarters")
lines(hpfilter(cafe, type = "lambda")$trend, col = "red", lwd = 2)
legend(1985, 8000, c("Original Serie", "Trend - HP Filter"), col = c("black", "red"), lwd = c(1,2), bty = "n")
plot(hpfilter(cafe, type = "lambda")$cycle,  xlab = "Time", ylab = "Cycle Component")

```

